---
title: "predictive-machine-learning-project"
author: "Hollismc"
date: "2/15/2022"
output: github_document
```{r}
#install.packages("credentials")
library("credentials")
credentials::set_github_pat(force_new = TRUE)
```

#Download libraries
```{r}
#install.packages(gbm)
library(knitr)
library(caret)
library(kernlab)
library(randomForest)
library(rattle)
library(gbm)
```
In order to predict the "manner in which they did the exercise", we used the "classe" variable as written in the project instructions. We first read the data into R with the code below, and then took a look at it.
```{r}
        #if(!file.exists("./data")) 
        #fileurl1 = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
        #fileurl2 = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
        #download.file(fileurl1, destfile = "./data/training.csv", method = "curl")
        #download.file(fileurl2, destfile = "./data/testing.csv", method = "curl")
        testing = read.csv("./data/testing.csv")
        training = read.csv("./data/training.csv")
head(training)
str(training)
```
#Preprocessing:
# clean up data set and exclude near zero data 
```{r}
nzvcol <- nearZeroVar(training)
training <- training[, -nzvcol]
```
# remove first 7 columns (not needed for predicting)
```{r}
train_1 <- training[,8:length(colnames(training))]
test_1 <- testing[,8:length(colnames(testing))]
```
#remove columns with NAS
```{r}
trainfinal <- train_1[,colSums(is.na(train_1)) == 0]
testfinal <- test_1[,colSums(is.na(test_1)) == 0]
dim(trainfinal)
```
#set seed for reproducibility
```{r}
set.seed(5678)
str(trainfinal)
```
#partition data into validation set
```{r}
train_2 <- createDataPartition(trainfinal$classe, p = 0.70, list = FALSE)
training <- trainfinal[train_2,]
validation <- trainfinal[-train_2,]
```
# model fit for random forest
```{r}
model <- trainControl(method = "cv", 5)
modelfit <- train(classe ~ ., data = training, method = "rf", trControl = model, ntree = 251)
print(modelfit)
```
#Predict
```{r}
predictval <- predict(modelfit, validation)
confuse <- confusionMatrix(predictval, factor(validation$classe))
confuse
```
#Plot
```{r}
plot(modelfit)
```
#Decision Trees
```{r}
modeltrees <- train(classe ~., data=training, method="rpart", trControl = model, tuneLength =5)
fancyRpartPlot(modeltrees$finalModel)
```
#Prediction
```{r}
predicttrees <- predict(modeltrees, validation)
confusetrees <- confusionMatrix(predicttrees, factor(validation$classe))
confusetrees
```
#Generalized Boosted Model
```{r}
modelGBM <- train(classe ~., data = training, method = "gbm", trControl = model, tuneLength = 5, 
                  verbose =FALSE)
predGBM <- predict(modelGBM, validation)
cmGbm <- confusionMatrix(predGBM,factor(validation$classe))
```
#Plot
```{r}
plot(modelGBM)
```
#Random Forest to Predict test Results
#Quiz
```{r}
pred <- predict(modelfit, testing)
print(pred)
```







